Analysis of NLA17 Dissolved Gas Indicator

### 2020-12-16 scripts/dgIndicatorAnalysis.Rmd
Added Random Semivariograms code chunk to dgIndicatorAnalysis

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot
library(ncf) # ncf::correlogram

@@

>>>>

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot
library(ncf) # ncf::correlogram
library(lattice) # xyplot

@@

## Random Semivariograms
Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.  Downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.  An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.  Then pH values are randomly sampled and assigned to the coordinates so that a random semivariogram can be calculated.  That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram.

The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)
# also includ WSA9, WSA9_NAME feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, wsa9, wsa9_name, onoffnet, nitrate_n_result, chla_result_vol, ntl_lab_result, ph_result, ptl_lab_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, state, size_cat, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & wsa9 != "AK")

names(dg_1)

#filter in View returns 514 nitrate_n_result equal to zero
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
  mutate(log_ntl = round(log(ntl_lab_result),3)) %>% 
  mutate(log_ptl = round(log(ptl_lab_result),3)) %>% 
  mutate(log_chla = round(log(chla_result_vol),1)) %>% 
  mutate(log_si = round(log(silica_result),2)) %>% 
  mutate(log_ca = round(log(calcium_result),2)) %>% 
  mutate(ph_result = round(ph_result,2)) %>% 
  mutate(nitrate_n_b = case_when(
    nitrate_n_result <= 0.0001 ~ "low",
    nitrate_n_result > 0.0001 ~ "high"))

dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
summary(dg_1$nitrate_n_b)

# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

st_crs(dg_sf1)

z_ph_sf1 = variogram(ph_result~1, dg_sf1)
plot(z_ph_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")

print(xyplot(gamma ~ dist, z_ph_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
))

z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
plot(variogram(ph_result~1, dg_sf1, cressie = TRUE))

print(xyplot(gamma ~ dist, z_cressie_ph, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1, cressie = TRUE)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'semivariance'
))

z_do_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
plot(z_do_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
plot(z_do_sf1, ylab = "NLA Surface DO Semivariance")
